\classheader{2018-07-09}
\subsection*{Complex Roots of the Characteristic Equation}
Back to the constant coefficients case:
\begin{equation*}
	ay'' + by= + cy = 0 \tag{$\star$}
\end{equation*}
Let $a = 1 = c$, $b = 0$. Here $y'' + y = 0$ has characteristic equation $\Gamma^2 + 1 = 0$ (no real solutions).\\
But we know $y_1(t) = \cos t$, $y_2(t) = \sin t$. Solve $y'' + y = 0$
\begin{equation*}
	\tcbhighmath[drop fuzzy shadow]{y(t) = c_1 \cos t + c_2 \sin t}
\end{equation*}
\redhline
\begin{center}
	\textbf{Question:} How can we get that from the characteristic equation?
\end{center}
\redhline\\
First, $r^2 + 1 = 0$ does have 2 solutions: $r = \pm i$\\
Sticking to the exponential there, $y_1(t) = e^{it}$ and $y_2(t) = e^{-it}$ are the two solutions.\\\\
\textbf{Recall Euler's Formula}
\begin{equation*}
	e^{(a + ib)t} = e^{at}(\cos bt + i \sin bt)
\end{equation*}
\redhline
\begin{center}
	\textbf{Question:} Can we construct real solutions from there?
\end{center}
Suppose an ODE has characteristics equation.
\begin{equation*}
	a\Gamma^2 + b \Gamma + c = 0, \quad \Gamma = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} = \lambda \pm \mu i, \quad \mu \neq 0 
\end{equation*}
Writing 2 exponential solutions.
\begin{equation*}
	\begin{split}
			y_1(t) & = e^{(\lambda + \mu i)t}\\
			& = e^{\lambda t}(\cos \mu t + i\sin \mu t)
	\end{split}
	\quad \quad \quad 
	\begin{split}
			y_1(t) & = e^{(\lambda - \mu i)t}\\
			& = e^{\lambda t}(\cos \mu t - i\sin \mu t)
	\end{split}
\end{equation*}
We see they are not real. But real ODE must be real solutions!\\
The \underline{superposition}, any linear combination of $y_1$, $y_2$ is also a solution:
\begin{align*}
	\text{Hence } \quad \frac{1}{2}(y_1(t) + y_2(t))  & = \frac{1}{2} (e^{\lambda t} \cos \mu t + i \sin \mu t) + e^{\lambda t} \cos \mu t + i \sin \mu t\\
	& = e^{\lambda t} \cos \mu t \quad \text{is a solution}
\end{align*}
\begin{align*}
	\text{Hence } \quad \frac{1}{2i}(y_1(t) - y_2(t))  & = \frac{1}{2i} (e^{\lambda t} \cos \mu t + i \sin \mu t) - e^{\lambda t} \cos \mu t + i \sin \mu t\\ & = e^{\lambda t} \sin \mu t \quad \text{is a solution}
\end{align*}
lets call then $u(t) = e^{\lambda t} \cos \mu t$, \quad $v(t)=e^{\lambda t} \sin \mu t$\\
There a 2 real solutions (the real and imaginary parts of the original operator solutions.)\\
\underline{And} since $W(u,v) = $ calculate this $= \mu e^{2 \lambda t} \neq 0$ anywhere as long as $\mu \neq 0$, the fund set of solutions is.
\begin{align*}
	y(t) & = c_1 e^{\lambda t} \cos \mu t + c_2 e^{\lambda t} \sin \mu t \\
	& = e^{\lambda t}(c_1 \cos \mu t + c_2 \sin \mu t)
\end{align*}
\begin{example-N}
	$y'' + y = 0$. Roots of $\Gamma^2 + 1 = 0$ are $\Gamma = \pm i$. Here $\lambda = 0$, $\mu = 1$.
	\begin{equation*}
		y(t) = e^{0t}(c_1 \cos 1t + c_2 \sin 1t) = \boxed{c_1 \cos t + c_2 \sin t}
	\end{equation*}
\end{example-N}
\begin{example-N}
	Solve the IVP $y'' + 4y' + 13y = 0$, $y(0) = 2$, $y'(0) = 7$\\
	\underline{Solution}: The characteristic equation is $\Gamma^2 + 4\Gamma + 13 = 0$, and $\Gamma. = \frac{-4 \pm \sqrt{16 - 52}}{2} = -2 \pm 3i$\\\\
	Here, find set of solutions is 
	\begin{equation*}
		y(t) = e^{-2t}(c_1 \cos 3t + c_2 \sin 3t)
	\end{equation*}
	For a particular solution
	\begin{align*}
		y(0) & = e^{-2(0)}(c_1 \cos 3(0) + c_2 \sin 3(0)) = \boxed{2 = c_1}\\
		y'(0) & = -2e^{-2t}(c_1 \cos 3t + c_2 \sin 3t) + e^{-2t}(-6 \sin 3t + 3c_2 cos 3t) \Bigg|_{t=0} \\
		& = -4 + 3c_2 = 7 \quad c_2 = \frac{11}{3}
	\end{align*}
	\begin{equation*}
		\boxed{y(t) = e^{-2t}(2 \cos 3t + \frac{11}{3} \sin 3t)}
	\end{equation*}
\end{example-N}
\redhline\\
\underline{\textbf{Facts}} Given $ay'' + by' + cy = 0$ and its characteristic equation $a\Gamma^2 + b\Gamma + c = 0$ with roots $\Gamma_1$, $\Gamma_2$:
\begin{enumerate}[label=\protect\circled{\Roman*}]
	\item $\Gamma_1 \neq \Gamma_2$, where real roots
	\begin{equation*}
		y(t) = c_1 e^{\Gamma_1 t} + c_2 e^{\Gamma_2 t} \tag{real roots}
	\end{equation*}
	\item $\Gamma_1 = \lambda + i\mu \neq \lambda - i\mu = \Gamma_2$, complex roots
	\begin{equation*}
		y(t) = e^{\lambda t} (c_1 \cos \mu t + c_2 \sin \mu t) \tag{complex roots}
	\end{equation*}
	\item $\Gamma_1 = \Gamma_2 = \Gamma$, equal roots
	\begin{equation*}
		y(t) = c_1 e^{\Gamma t} + c_2 e^{\Gamma t} = (c_1 + c_2)e^{\Gamma t} = K e^{\Gamma t} \tag{equal roots}
	\end{equation*}
	Is only 1 solution. We will need another linear independent one to construct a fund set of solutions.
\end{enumerate}